---
title: "Data sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EstZoonoticTB)
```


# Bovine tuberculosis (TB) incidence estimates

## Humans

- [ ] Bovine TB in humans by country (if available). 
    - Source from [this systematic review](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4816377/)
## Animals

- [ ] Bovine TB in animals.
  - Available as an indicator
  - Stratified by wild and domesticated

# Tuberculosis (TB) incidence

* Sourced from the WHO using WHO data
* Extract TB incidence, proportion in children, proportion HIV positive, and proportion drug resistant.
* Data sourced using `EstZoonoticTB::get_tb_data()`. See `?EstZoonoticTB::get_tb_data()` for documentation.

# Animal reservoir populations

## Domesticated

* Cattle
    - [Source](http://www.fao.org/faostat/en/#search/Cattle)
* Other
    - Need to know what animals can have TB.
    - [Source](http://www.fao.org/faostat/en/?#data/QA)
    
## Wildlife

* What wild animals?
* [Source](http://www.fao.org/faostat/en/?#data/QA)


## Demographics

- Population data by country.
   - Available as rural and urban split.
   - [Source](http://www.fao.org/faostat/en/#search/Cattle)
   


- [x] TB incidence by country.
    - Estimates via [getTBinR](https://github.com/seabbs/getTBinR)