---
title: "Data sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EstZoonoticTB)
```


# Bovine tuberculosis (TB) incidence estimates

## Humans

- [ ] Bovine TB in humans by country (if available). 
    - Source from [this systematic review](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4816377/)
    
## Animals

* Sourced from the [OIE](https://www.oie.int) via personal communication
* Data cleaning details can be found in [/data-raw/zoonotic_tb_animals.R](https://github.com/seabbs/EstZoonoticTB/blob/master/data-raw/zoonotic_tb_animals.R)
* Data available for 2018 split into half years.
* Data is available for domesticated and wild animals - stratified into present, limited, suspected, limited + suspected, not present and missing. 
* Incidence or incidence rates are not available in this dataset.
* Data quality is low.
* Last recorded status is also available but this field has currently been excluded due to its poor quality. 

```{r}
EstZoonoticTB::zoonotic_tb_animals
```

# Tuberculosis (TB) incidence

* Sourced from the [WHO](http://www.who.int/tb/country/data/download/en/) using [`{getTBinR}`](https://www.samabbott.co.uk/getTBinR/index.html).
* Extracted TB incidence, TB incidence rates (+ CI's), and proportion HIV positive (+ CI's).
* Data sourced using `EstZoonoticTB::tb_data()`. See `?EstZoonoticTB::tb_data()` for documentation.


```{r}
EstZoonoticTB::tb_data()
```


## Demographics

* Sourced from the [FAO](http://www.fao.org/faostat/en/?#data/OA). 
* Data cleaning can be found in [/data-raw/demographics.R](https://github.com/seabbs/EstZoonoticTB/blob/master/data-raw/demographoics.R).
* Population estimates are available by country from 1950 until 2018. 
* The proportion of each country that is rural is also available.


```{r}
EstZoonoticTB::demographics
```
   
   
# Animal Demographics

## Domesticated

* Sourced from the [FAO](http://www.fao.org/faostat/en/?#data/OA). 
* Data cleaning can be found in [/data-raw/animal_demographics.R](https://github.com/seabbs/EstZoonoticTB/blob/master/data-raw/animal_demographics.R).
* Animals included:
  * Cattle
* Additional animals that are vecotors of Zoonotic TB could also be included as needed/thought appropriate

```{r}
EstZoonoticTB::animal_demographics
```
   
