---
title: "Data sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EstZoonoticTB)
```


# Bovine tuberculosis (TB) incidence estimates

## Humans

- [ ] Bovine TB in humans by country (if available). 
    - Source from [this systematic review](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4816377/)
## Animals

- [ ] Bovine TB in animals.
  - Available as an indicator
  - Stratified by wild and domesticated

# Tuberculosis (TB) incidence

* Sourced from the [WHO](http://www.who.int/tb/country/data/download/en/) using [`{getTBinR}`](https://www.samabbott.co.uk/getTBinR/index.html).
* Extracted TB incidence, TB incidence rates (+ CI's), and proportion HIV positive (+ CI's).
* Data sourced using `EstZoonoticTB::tb_data()`. See `?EstZoonoticTB::tb_data()` for documentation.
* Countries are excluded if incidence was below 100 or if incidence rates were below 1.


```{r}
EstZoonoticTB::tb_data()
```

# Animal reservoir populations

## Domesticated

* Cattle
    - [Source](http://www.fao.org/faostat/en/#search/Cattle)
* Other
    - Need to know what animals can have TB.
    - [Source](http://www.fao.org/faostat/en/?#data/QA)
    
## Wildlife

* What wild animals?
* [Source](http://www.fao.org/faostat/en/?#data/QA)


## Demographics

* Sourced from the [FAO](http://www.fao.org/faostat/en/?#data/OA). 
* Data cleaning can be found in [/data-raw/demographics.R](https://github.com/seabbs/EstZoonoticTB/blob/master/data-raw/cleaning.R).
* Population estimates are available by country from 1950 until 2018. 
* The proportion of each country that is rural is also available.


```{r}
EstZoonoticTB::demographics
```
   

