---
title: "Data linkage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data linkage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette outlines the data linkage process used to link various packaged - and potentially user supplied - datasets. Further details are provided in the linked function documentation.


## Linkage

The datasets outlined in the [data sources](https://www.samabbott.co.uk/EstZoonoticTB/articles/data-sources.html) vignette are linked using country code (ISO3) where available and country name (using an automated approach with manual checks). This results in a single dataset which includes data on zoonotic TB and various correlated variables. Data completeness is low with some datasets spanning multiple decades (from the 1950's) and some representing only a single year of data. Default datasets can be updated by passing datasets in the required format to arguments of `link_data` (see [`?link_data`](https://www.samabbott.co.uk/EstZoonoticTB/reference/link_data.html) for details).

### Code

```{r}
EstZoonoticTB::link_data
```

### Data

```{r}
all_data <- EstZoonoticTB::link_data(verbose = TRUE)

all_data
```

### Latest data

As many of the packaged datasets are highly missing a more useful dataset for analysis may be one that includes the most recent estimate in each country for each variable. This can be extracted using `get_latest_combined_data`.

### Code

```{r}
EstZoonoticTB::get_latest_combined_data


EstZoonoticTB::get_latest_data
```

### Data

```{r}
latest_data <- EstZoonoticTB::get_latest_combined_data(all_data)

latest_data
```